# Predicting learning

```{r}
library(ggplot2)
library(broom)
options(scipen=999, digits=3)

source('../scripts/raw2tidy.R')
pilotsToInclude <- c(4, 5)
dataToInclude <- c('decision', 'learning', 'demographics')
x <- prepareAll(pilotsToInclude)

source('../scripts/lib/combinePilots.R')
x <- combinePilots(x, paste0('p', pilotsToInclude), dataToInclude)

source('../scripts/addFeatures.R')
x <- addFeatures(x)
```

```{r}
attach(x)
```

## Predicting (Boolean) acquisition and reversal
We regress on model-free and model-based measures of decision-making separately, as well as on age and the degree of statistical training. Only age turns out to be a good predictor; even when removed, decision-making measures don't predict acquisition and reversal well.

```{r}
y <- merge(learningFeatures, decisionFeatures) %>% merge(demographics) %>% 
    select(-LL, -iterations, -message, -income, -ends_with('.A'), -ends_with('.B')) %>%
    mutate(status=as.factor(status), acquired=learned.Acq2, reversed=learned.Rev3)
y %>% glimpse()
```

```{r}
# logreg of trial properties predicting acquisition
model.acquired <- glm(acquired ~ risk_prop + ambig_prop + age + stats_training, family = binomial(link = 'logit'), 
                      data = y)
# summary(model.choice)
tidy(model.acquired)
glance(model.acquired)
```

```{r}
model.acquired.mb <- glm(acquired ~ alpha + beta + age + stats_training, family = binomial(link = 'logit'), 
                      data = y)
# summary(model.choice)
tidy(model.acquired.mb)
glance(model.acquired.mb)
```

```{r}
# logreg of trial properties predicting acquisition
model.reversed <- glm(reversed ~ risk_prop + ambig_prop + age + stats_training, family = binomial(link = 'logit'), 
                      data = y)
# summary(model.choice)
tidy(model.reversed)
glance(model.reversed)
```

```{r}
model.reversed.mb <- glm(reversed ~ alpha + beta + age + stats_training, family = binomial(link = 'logit'), 
                      data = y)
# summary(model.choice)
tidy(model.reversed.mb)
glance(model.reversed.mb)
```

### Is age actually a good predictor, or is the result driven by outliers?

```{r}
options(repr.plot.width=4, repr.plot.height=2)
ggplot(y, aes(x=age, fill=acquired)) + geom_histogram(binwidth=1.5, alpha=.5, position="identity") + theme_bw()
ggplot(y, aes(x=age, fill=acquired)) + geom_density(alpha=.5) + theme_bw()
```

```{r}
options(repr.plot.width=4, repr.plot.height=2)
ggplot(y, aes(x=age, fill=reversed)) + geom_histogram(binwidth=1.5, alpha=.5, position="identity") + theme_bw()
ggplot(y, aes(x=age, fill=reversed)) + geom_density(alpha=.5) + theme_bw()
```

```{r}
options(repr.plot.width=4, repr.plot.height=2)
jitter_and_model <- list(
    geom_jitter(height=.05, alpha=.5),
    stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE),
    scale_y_continuous(name=NULL, breaks=c(0, 1), labels=c("0"='No', "1"='Yes'))
)
ggplot(y, aes(x = age, y = as.numeric(acquired))) + jitter_and_model + ggtitle('Did the subject successfully acquire?')
ggplot(y, aes(x = age, y = as.numeric(reversed))) + jitter_and_model + ggtitle('Did the subject successfully reverse?')
```