# Basic risk & ambiguity analysis
Including pilots 2, 4, and 5, since their R&A task was the same.

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

options(scipen=999, digits=3)

source('../scripts/raw2tidy.R')
pilotsToInclude <- c(2, 4, 5)
dataToInclude <- c("decision")
x <- prepareAll(pilotsToInclude)

source('../scripts/lib/combinePilots.R')
x <- combinePilots(x, paste0('p', pilotsToInclude), dataToInclude)
```

```{r}
source('../scripts/addFeatures.R')
x <- addFeatures(x)
```

```{r}
attach(x)
```

```{r}
x$decisionFeatures
```

### Display choice curves

```{r}
# Re-shape data into a longer format for this purpose
rna_l <- decision %>% mutate(trialtype = ifelse(al == 0,
                                           "probability",
                                           "ambiguity"),
                        quantity = ifelse(al == 0, p, al)) %>% 
                 group_by(ID, trialtype, val, quantity) %>% 
                 summarize(proportion = mean(choice))

rna_l$quantity[rna_l$quantity == .24] <- .25
head(rna_l)
```

## Choice curves of all subjects in one graph

```{r}
options(repr.plot.width=8, repr.plot.height=8)
subject_all_g = ggplot(rna_l, aes(x = val, y = proportion, color = as.factor(ID))) + geom_line(position=position_dodge(width=1.5)) + 
    facet_grid(quantity ~ trialtype) + geom_point() + xlab("Lottery payoff ($)") + ylab("Proportion of lotteries chosen") + 
                            ggtitle("Choice curve, by probability / ambiguity level") + theme_bw()
subject_all_g
```

```{r}
num_subjects <- length(unique(rna_l$ID))
```

```{r}
options(repr.plot.width = 8, repr.plot.height = 1 * num_subjects)

subject_all_g2 = ggplot(rna_l, aes(x = val, y = proportion, color = as.factor(quantity))) + geom_line(position=position_dodge(width=0.5)) + 
    facet_grid(ID ~ trialtype) + geom_point() + xlab("Lottery payoff ($)") + ylab("Proportion of lotteries chosen") + 
                            ggtitle("Choice curve, by subject") + theme_bw()
subject_all_g2
```

## Compare R fits to Matlab fits

```{r}
matlab <- read.csv('../data/processed/decision_matlab.csv')  %>% rename(ID = id)
head(matlab)
head(decisionFeatures)
```

```{r}
comparison <- merge(decisionFeatures %>% select(-message),
                    matlab %>% select(ID, alpha, beta, gamma, exitflag, LL), 
                    by = 'ID', suffixes=c('.R', '.matlab'))  %>% unique()
```

It seems that R has log-likelihood results that are closer to 0 more often than Matlab.

```{r}
comparison  %>% group_by(ID) %>% 
    mutate(alpha_d = alpha.R - alpha.matlab, 
           beta_d = beta.R - beta.matlab, 
           gamma_d = gamma.R - gamma.matlab,
           LL_d = LL.R - LL.matlab)  %>% 
    filter(abs(alpha_d) > 0.03 | abs(beta_d) > 0.03)
```

## Testing stochastic dominance
It is a violation of stochastic dominance if someone chooses an uncertain option with a given payoff instead of a certain option with that payoff. This should be true regardless of the level of ambiguity or risk. At best, the violation indicates irrationality; at worst, it indicates a failure to understand the task.

```{r}
rna_l %>% filter(val==5, proportion > 0)
rna_l %>% filter(val==5, proportion > 0) %>% group_by(ID) %>% summarize(violations=sum(proportion))
```

Should the violators be excluded? Should there be a tolerance for one or two stochastic-dominance violating decisions?

## Visualizing the decision matrix

```{r}
decision_matrix <- rna_l %>% unite(trialq, trialtype, quantity)
head(decision_matrix)
```

```{r}
adjust_trial_label <- function(str) {
    str <- gsub('^([ap]).+_(0\\.)?(.+)', '\\1.\\3', str)
    str <- gsub('.1$', '.full', str)
}
```

The left side of each graph denotes the choices under ambiguity; the right choices under risk. Both increase from left to right.

*How to tell the choices are consistent?* One, the closer to the red plus, the darker the plot should be. Two, the gradient should be even, without any sudden jumps back and forth. Three, the bottom horizontal strip that denotes the value of $5 should be all dark, lest there be a violation of stochastic dominance.

```{r}
options(repr.plot.width = 10, repr.plot.height = 8)
ggplot(data=decision_matrix, aes(x=adjust_trial_label(trialq), y=as.factor(val), fill=proportion)) + 
    geom_tile() + 
    facet_wrap(~ ID, shrink = TRUE) + 
    geom_vline(aes(xintercept = 4.5)) + geom_hline(aes(yintercept = 1.5)) +
    geom_point(x = 4.5, y=.5, size = 4, color = "red", shape='+') +
    theme(axis.text.x = element_text(angle=45, vjust=0.5)) + 
    ggtitle('Decision matrix for Pilots 4 and 5') + ylab('Payoff ($)') + 
    xlab('Trial kind - [a]mbiguous (% occluded) or [p]robabilistic (% win chance)')
```